Detection and partial characterisation of stellar multiplicity with Gaia

### 2018-08-21 article/ms.tex
filler figures

\documentclass[twocolumn]{aastex61}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{color}

@@

@@

\begin{abstract}
The properties and occurrence rate of stellar multiples (e.g., binaries, triples) 
underpins much of astrophysics. 
%Precisely measuring these quantities is challenging and observationally demanding. 
Here we use \Gaia\ data to detect and partially 
characterise \NumberOfStellarMultiples\ stellar multiple systems based on excess 
jitter in astrometry and radial velocity.
Comparisons with literature surveys on stellar multiplicity demonstrate that we 
reliably detect binary systems with orbital periods up to $\approx$3.5\,yr. 
We use eclipsing binary systems to calibrate excess astrometric jitter with
tangential motion, allowing us to estimate inclination angles and directly constrain
companion masses for millions of systems.
\todo{A comment on stellar multiplicity with stellar metallicity.}
\todo{A comment on stellar multiplicity in the field relative to clusters.}
\todo{A comment on systems with stellar remnants (e.g., NS/BHs).}
\end{abstract}

@@

Higher order stellar systems describes any stars that have at least one stellar 
companion (e.g., binaries, trinaries). The presence of stellar companions
introduces a number of complications to the inference of stellar and galactic
properties, but these factors are often ignored.

The \Gaia\ space telescope provides exquisite astrometry, photometry, and radial
velocity measurements over many years for million of point stars in our galaxy.
%These data provides an excellent opportunity to self-calibrate the errors in those
%measurements. 
When a point source has an unresolved stellar companion within some range of orbital
and stellar parameters, it is expected that there will be an anomalous excess in 
these measured properties relative to a single star of a similar type. This is
nothing new: astronomers have used radial velocity \citep{Butler:1996} and 
astrometric variations \citep{Muterspaugh:2010} and photometry \citep{someone} to infer the 
presence of stellar and exoplanet companions forever. However, the opportunity
presented by \Gaia\ is to infer the presence of stellar companions for millions
of point sources, across both hemispheres. No ground- or space-based instrument
has ever provided such an opportunity.

In this work we make use of photometry and astrometry (including radial motion)
measurements in the second \Gaia\ data release to infer the presence of stellar
companions for millions of point sources, and provide some basic characterisations
of these systems. In Section \ref{sec:method} we describe our methods, and in
Section \ref{sec:results} we outline our results in context of other stellar
multiplicity surveys. We include comparisons of binary properties between
clusters and the field, as a function of stellar properties (e.g., stellar
metallicity), and highlight some particularly noteworthy systems that would
benefit from immediate follow-up spectroscopic observations.

@@

There are many ways that binaries can be reliably identified using \Gaia\
data, including radial velocity measurements, photometry, and astrometry.
These detection methods are sensitive to binary systems with different 
properties. Here we describe the absence of reported radial velocity in
\Gaia\ data can be used to deduce stellar multiples, before we describe
our model that incorporates measurements of radial velocity, astrometry, 
and photometry, to detect and characterise stellar multiples.

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{../figures/sb2_rvs_completeness.pdf}
    \caption{Fraction of \Gaia\ sources with reported radial velocities
		     as a function of source properties. Within the adopted source
		     parameter ranges (gray; see Section \ref{sec:sb2_methods})
		     the radial velocity completeness is approximately flat.
		     We flag sources within this range that do not have reported 
		     radial velocities to be candidate SB2 systems.}
    \label{fig:sb2_rvs_completeness}
\end{figure*}

\begin{figure}
	\includegraphics[width=0.4\textwidth]{../figures/sb2_histograms.pdf}
    \caption{Distribution of 
    			(a) astrometric unit weight error (see Eq. \ref{eq:astrometric_unit_weight_error}),
			%$u = \left(\chi_{al}^2/(N_{al} - 5)\right)^{1/2}$
			%		where $\chi_{al}^2$ and $N_{al}$ is the astrometric goodness-of-fit and number 
			%		of observations in the along-scan direction, respectively (columns 
			%			\texttt{astrometric\_chi2\_al} and \texttt{astrometric\_n\_obs\_al}), 
				(b) photometric \bp\ - \rp\ excess factor, 				
	 				and 
				(c) photometric \rp\ variability (see Eq. \ref{eq:photometric_variability})
			of candidate SB2 type systems relative to a control sample of the same size.}
    \label{fig:sb2_histograms}
\end{figure}

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{../figures/sb2_sky_structure.pdf}
    \caption{Fraction of sources (per sky bin) without reported radial velocities
    		 for all sources with $G \lesssim 13$ (top) and sources in the ranges
		 	 specified by Eqs \ref{eq:sb2_ranges} (bottom), where we assert that a missing
			 radial velocity measurement signifies a likely SB2 candidate . The color scale is 
			 arbitrarily set to highlight structure, where black indicates a higher
			 fraction of sources do not have reported radial velocities. Crowding
			 in the galactic plane likely results in some sources not having radial
			 velocities reported. The large scale structure visible in both axes is
			 a combined effect of the initial \Gaia\ source list, the scanning law,
			 and star forming regions (i.e., where emission in the \ion{Ca}{2} 
			 triplet likely causes issues for radial velocity determination).}
    \label{fig:sb2_sky_structure}
\end{figure*}

\subsection{SB1 Systems: Single-lined spectroscopic binaries}
\label{sec:sb1_methods}

Radial velocity measurements are not available for individual epochs in the 
second \Gaia\ data release \citep{Lindegren:2018,Cropper:2018}, but a point estimate
and error in radial velocity is available for 7,224,631 sources \cite{Cropper:2018}.
% TODO: That's the value from the paper, but I think the true value is lower. Check.
The reported radial velocity error ($\GaiaRVE$; column name \texttt{radial\_velocity\_error}) 
is a function of the number of transits $N$ (i.e., the number of observations; 
represented by column \texttt{rv\_nb\_transits}) and the standard deviation among 
those $N$ measurements of radial velocity. This allows us to calculate the standard
deviation in radial velocity for each source, independent of the number of measurements,
which we will refer to as the \emph{radial velocity jitter} $\RVJitter$,
\begin{eqnarray}
\RVJitter = \sqrt{\frac{2N}{\pi}}\GaiaRVE \quad .
\end{eqnarray}

For a single star (without a significant mass companion) the source radial velocity
jitter represents the minimum uncertainty with which \Gaia\ can measure 
radial velocity for a star of that colour, apparent brightness, and absolute brightness.
However for stars with stellar companions and no evidence of double lines or binarity in the
cross-correlation-function, so-called SB1 type systems \citep{someone}, the
radial velocity jitter is a function of the orbital parameters and observation
epochs. We assume that the distribution of radial velocity jitter in the second
\Gaia\ data release is a mixture of two populations:

\begin{enumerate}
\item \emph{single-star systems}, where the radial velocity jitter represents the
      minimum radial velocity noise that \Gaia\ can measure for a star of that
      \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude;

\item \emph{higher-order star systems}, where the radial velocity jitter is
      significantly above that for single-star systems of a similar colour, 
      apparent magnitude, and absolute magnitude.
\end{enumerate}

The mean radial velocity jitter of a single star will vary depending on the 
\bp\ - \rp\ colour of the source and the apparent magnitude. Similarly, the mean
radial velocity jitter of stellar \emph{multiples} will depend on the source 
colour and its position in the Hertzsprung-Russell diagram. For these reasons,
we must consider the radial velocity jitter of a source in context with other
sources that have a similar \bp\ - \rp\ colour, a similar apparent \rp\ magnitude,
and a similar absolute \rp\ magnitude. In principle if we knew the underlying
distribution and properties of stellar multiples then we could model how different
binary (and higher order) systems would be observable across the Hertszprung-Russell
diagram.

We chose to adopt a non-parametric model for stellar multiplicity across the
Hertsprung-Russell diagram because we do not know the distribution of properties
of stellar multiples. Specifically, for each source we take the following steps:

\begin{enumerate}

	\item Select sources with similar \bp\ - \rp\ colours, a similar apparent \rp\
		  magnitude, and a similar absolute \rp\ magnitude.

	\item We fit a two-component mixture model to the distribution of radial
		  velocity jitter for the subset of similar sources selected in Step 1.

	\item Evaluate whether the source of interest is more consistent with being
		  drawn from the single star component (the normal distribution) or the
		  stellar multiple component (the log-normal distribution).

	\item For sources that are more consistent with being drawn from the stellar
		  multiple component, provide a partial characterisation of the system by
		  calculating the excess radial velocity jitter (above what is expected 
		  for a single star with similar observed properties), among other 
		  quantities.

@@

We illustrate this procedure in Figure \ref{fig:npm_schematic}, where we demonstrate
how the distribution of source properties (e.g., radial velocity jitter) varies across
the Hertsprung-Russell diagram. While this model is flexible enough to describe varying
properties of single stars and stellar multiples, there are a number of explicit 
assumptions that go into this model:

\todo{assumptions}

@@

\subsubsection*{Step 1: Selecting similar sources}

We construct a $k$-dimensional tree \citep[k-d tree;][]{} for all sources in the
dimensions of \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude.
When doing so we scale the dimensions such that 0.1\,mags in \bp\ - \rp\ colour is
approximately equal to 1 magnitude in apparent \rp\ magnitude and 1 magnitude in 
absolute \rp\ magnitude. We use this k-d tree for constructing the `ball' of similar
stars for every source.

We enforce a number of constraints when querying for similar stars. For each source
we require the `ball' in multi-dimensional space to include at least 128 sources.
We further require that the radius of the ball extends to at least 0.05\,mags in 
\bp\ - \rp\ colour, 0.5\,mags in apparent \rp\ magnitude, and 0.5\,mags in absolute
\rp\ magnitude. If the initial query of 128 points does not extend wide enough to
meet our constraints on minimum radius, we extend the radius (linearly in each scaled)
dimension until the minimum constraints on radii are met. Our minimum radii constraints
imply that there will be many (e.g., $>>128$) points returned in dense regions of the
Hertsprung-Russell diagram. In these situations we then subsample the selected points
to return a random 1024 sources in any ball, while enforcing our constraints on
minimum radii. As a result, the number of `similar' sources selected will range from
a minimum of 128 to a maximum of 1024, and the minimum radii constraints on various
dimensions are always maintained.

\subsubsection*{Step 2: Optimizing the two-component mixture}

We fit a two-component mixture model to the distribution of radial velocity jitter for
all similar sources selected through Step 1. This mixture model includes a normal
distribution to represent the intrinsic jitter for single stars, and a log-normal
distribution to represent stellar multiples. Specifically the model parameters include
a mixing weight parameter $\theta$ that is constrained between $(0, 1)$, the mean 
$\mu_{s}$ and standard deviation $\sigma_{s}$ of the normal distribution,
and the mean $\mu_{m}$ and standard deviation $\sigma_{m}$ of the 
log-normal distribution. For brevity we

We assume a beta prior on $\theta$  such that $\theta \sim \mathcal{B}\left(5, 5\right)$
and we require $\sigma_{s}$ and $\sigma_{m}$ to be larger than zero.
Depending on the distribution of the radial velocity jitter, we often found this mixture
challenging to optimize because the log-normal distribution could easily account for 
most of the observed distribution, which tended the mixing weight to zero and implied
that all selected sources were stellar multiples. To avoid this situation we placed a
constraint on the mean of the log-normal distribution to ensure that the \emph{mode}
of the log-normal distribution is between $\mu_{s} + 1\sigma_{s}$ and
$\mu_{s} + 5\sigma_{s}$. The mode of the log-normal distribution is
$\exp\left(\mu_{m} - \sigma_{m}^2\right)$ such that
\begin{eqnarray}
	\mu_{s} + 5\sigma_{s} & > & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_lower_bound} \\
	\mu_{s} + 1\sigma_{s} & < & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_upper_bound} \quad .
\end{eqnarray}

However, we found that this constraint alone was insufficient. Often the mode would
be very close to the $\mu_{s} + 1\sigma_{s}$ lower bound, which implies that
the log-normal distribution is providing a considerable fraction of probability support
to sources with radial velocity jitter near the mean of the \emph{normal} distribution.
For this reason we chose to alter our lower bound constraint in that we also required
that at the mean of the normal distribution, the cumulative probability distribution
of the log-normal must be less than 10\% (let $q = 0.10$). The cumulative distribution 
function for the log-normal distribution is
\begin{eqnarray}
\textrm{CDF}(x|\mu_{m},\sigma_{m}) = \frac{1}{2} + \frac{1}{2}\textrm{erf}\left(\frac{\log{x} - \mu_{m}}{\sqrt{2}\sigma_{m}}\right) 
\end{eqnarray}

@@

We found these priors were sufficient to ensure numerical stability and efficiency.

Wherever possible we initialised the model parameters from nearby points (within the
ball selected by the k-d tree) where an optimised result already exists. This was primarily to minimise
computational cost; the initialisation did not have a noticeable impact on the
optimised result. We required the log probability and the gradient of the log
probability to reach machine precision before considering the optimisation
converged. After convergence, we queried to see if any points within the ball of
similar stars had not been optimized. If no result existed, we would next move to that
point. We refer to this process as \emph{swarm optimisation}, which we conducted in 
parallel, and illustrate in Figure \ref{fig:swarm_optimisation}.

@@

\subsection{Astrometric binaries} \label{sec:method_astrometric_binaries}

Many higher-order star systems are unresolved point sources in \Gaia, but the
stellar companion(s) may have sufficient mass to measurably affect the tangential 
velocity of the primary and produce a detectable jitter in the astrometric 
positions. There are many other sources of astrometric jitter, but here we show 
that the principle effect of astrometric excess noise is due to stellar 
companions. Note that due to the astrometry `degree of freedom bug' in the 
processing of the second \Gaia\ data release \citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, 
that resulted in 80\% of sources having zero astrometric excess noise,
we instead make use of the astrometric unit weight error $u$
\begin{eqnarray}
	u = \sqrt{\frac{\chi^2_{al}}{N_{obs,al} - 5}} \label{eq:astrometric_unit_weight_error}
\end{eqnarray}
\noindent{}where $\chi^2_{al}$ is the astrometric $\chi^2$ value
in the along-scan direction (column \texttt{astrometric\_chi2\_al}) and 
$N_{obs,al}$ is the number of astrometric observations in the along-scan
direction (column \texttt{astrometric\_n\_obs\_al}). The values of this
quantity can be interpreted similarly to how the astrometric excess noise
would be interpreted (e.g., large values are less consistent with a
single star model), although like the radial velocity and astrometric
jitter, it must be considered in context with stars of similar colour,
apparent magnitude (and preferably, absolute magnitude).

\subsection{Photometric binaries} \label{sec:method_photometric_binaries}

\todo{bp rp excess factor}

\begin{eqnarray}
	\textrm{Amp} = \log_{10}\left(\sqrt{N_{obs,al}}\frac{\sigma_{\langle{}f_G\rangle}}{\langle{}f_G\rangle}\right) \label{eq:photometric_variability}
\end{eqnarray}

\section{Results}

>>>>

\documentclass[twocolumn]{aastex61}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{color}
\usepackage{comment}

@@

\author{Rosemary Mardling}
\author{John Lattanzio}

@@

\begin{abstract}
The frequency and properties of stellar multiples (e.g., binaries, triples) underpin 
much of astrophysics, and are difficult to estimate.  
Here we use \Gaia\ data to detect and partially characterise \NumberOfStellarMultiples\
stellar multiple systems based on excess jitter in astrometry and radial velocity.
We reliably detect binary systems with orbital periods up to $\approx$3.5\,yr. % and systems with what kind of mass ratios, semi-amplitudes,  etc. 
We use eclipsing binary systems to calibrate excess astrometric jitter with
tangential motion, allowing us to estimate inclination angles and directly constrain
companion masses for millions of systems.
\todo{We find that the distribution of inclination angles is isotropic.}
\todo{A comment on stellar multiplicity with stellar metallicity.}
\todo{A comment on stellar multiplicity in the field relative to clusters.}
\todo{A comment on systems with stellar remnants (e.g., NS/BHs).}
\end{abstract}

@@

A higher order stellar system describes any star that has at least one stellar 
companion (e.g., binaries, trinaries). The presence of stellar companions
complicates inferences on stellar and galactic properties, but these factors are
often ignored because it is extremely challenging to detect stellar multiplicity
for an unresolved system, and even harder to separate the contributions from individual
sources.

The \Gaia\ space telescope provides exquisite astrometry, photometry, and radial
velocity measurements over many years for million of point sources in our galaxy
\citep{GaiaDR2}.
When a point source has an unresolved stellar companion within some range of orbital
and stellar parameters, it is expected that there will be an excess in motion jitter
relative to a single star of a similar type. This is nothing new: astronomers have 
used radial velocity \citep{Butler:1996} and astrometric variations \citep{Muterspaugh:2010}
to infer the presence of stellar and exoplanet companions for decades. However, 
the \Gaia\ data presents an opportunity to infer the presence of stellar companions
for millions of point sources across both hemispheres. No ground- or space-based 
instrument has ever provided such data.

\todo{Literature review on what we do and don't know about  the multiplicity fraction}

In this work we make use of astrometry (including radial motion) \todo{and photometry?}
from the second \Gaia\ data release to infer the presence of stellar companions 
for millions of point sources. We provide partial characterisations of these systems,
based on the data available for each source. In Section \ref{sec:method} we 
describe our methods, and in Section \ref{sec:results} we outline our results 
in context of other stellar multiplicity surveys. We include comparisons of 
binary properties between clusters and the field, as a function of stellar 
properties (e.g., stellar metallicity), and highlight some particularly 
noteworthy systems that would benefit from additional observations.

\section{Data} \label{sec:data}

There are many ways that binaries can be reliably identified using \Gaia\
data, including radial velocity measurements, astrometry, and photometry.
These data are sensitive to characterising stellar multiplicity in 
different ways. Here we describe the data that we use from the second
\Gaia\ data release \citep{Lindegren:2018} before defining our model.

%Here we define a model that we will use to detect and 
%characterise stellar multiples from radial velocity and astrometric data,
%before describing how the absence of \Gaia\ radial velocities can also be
%used to infer stellar multiplicity.

%Here we describe the absence of reported radial velocity in
%\Gaia\ data can be used to deduce stellar multiples, before we describe
%our model that incorporates measurements of radial velocity, astrometry, 
%and photometry, to detect and characterise stellar multiples.

\subsection{Radial velocities}
\label{sec:data_vrad}

Radial velocity measurements are not available for individual epochs in the 
second \Gaia\ data release \citep{Lindegren:2018,Cropper:2018}, but a point estimate
and an error in radial velocity is available for 7,224,631 sources \citep{Cropper:2018}.
% TODO: That's the value from the paper, but I think the true value is lower. Check.
The reported radial velocity error ($\GaiaRVE$; column name \texttt{radial\_velocity\_error}) 
is a function of the number of transits $N$ (i.e., the number of observations; 
given by column \texttt{rv\_nb\_transits}) and the standard deviation among 
those measurements. This allows us to calculate the standard deviation in radial 
velocity for each source, independent of the number of measurements,
which we will refer to as the \emph{radial velocity jitter} $\RVJitter$,
\begin{eqnarray}
\RVJitter = \sqrt{\frac{2N}{\pi}}\GaiaRVE \quad .
\end{eqnarray}

For a single star, or a star without a significant mass companion, the source radial
velocity jitter represents the minimum uncertainty with which \Gaia\ can measure 
radial velocity for a star of that colour, apparent magnitude, and absolute magnitude.
Stars with bluer colours have fewer absorption lines with deeper wings, and fainter stars
have on average lower signal-to-noise (S/N) ratios, both of which result in noisier radial
velocity measurements. The absolute magnitude is similarly important, as giant stars have
narrow absorption lines than main-sequence stars of the same temperature and colour. For stellar multiples, the source radial velocity jitter will be the quadrature sum of the expected jitter for the primary (if it were a single star) and a contribution based on the orbital parameters  and the observation epochs.

% todo: lack of RV measurements?

\subsection{Astrometry} \label{sec:data_astrometry}

Many higher-order star systems are unresolved point sources in \Gaia, but the
stellar companion(s) may have sufficient mass to measurably affect the tangential 
velocity of the primary and produce a detectable jitter in the astrometric 
position. There are many other sources of astrometric jitter, but here we show 
that the principle effect of astrometric excess noise is due to stellar 
companions. Note that due to the astrometry `degree of freedom bug' in the 
processing of the second \Gaia\ data release \citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, 
that resulted in 80\% of sources having zero astrometric excess noise,
we instead make use of the astrometric unit weight error $u$
\begin{eqnarray}
	u = \sqrt{\frac{\chi^2_{al}}{N_{obs,al} - 5}} \label{eq:astrometric_unit_weight_error}
\end{eqnarray}
\noindent{}where $\chi^2_{al}$ is the astrometric $\chi^2$ value
in the along-scan direction (column \texttt{astrometric\_chi2\_al}) and 
$N_{obs,al}$ is the number of astrometric observations in the along-scan
direction (column \texttt{astrometric\_n\_obs\_al}). The values of this
quantity can be interpreted similarly to how the astrometric excess noise
would be interpreted (e.g., large values are less consistent with a
single star model), although like the radial velocity and astrometric
jitter, it must be considered in context with stars of similar colour,
apparent magnitude, and preferably, absolute magnitude.

% todo: what ADQL queries did we do?
% what will our subset be?

@@

We assume that for sources with similar \bp\ - \rp\ colours, apparent \rp\ magnitudes, and absolute \rp\ magnitudes, the jitter distribution (in radial velocity or astrometry) is a mixture of two populations:

\begin{enumerate}
\item \emph{single-star systems}, where the jitter represents the minimum noise that 
	   \Gaia\ can measure for a star of that
      \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude;

\item \emph{higher-order star systems}, where the jitter is
      significantly above that for single-star systems of a similar colour, 
      apparent magnitude, and absolute magnitude.
\end{enumerate}

We also assume that the mean single-star jitter will change
depending on the \bp\ - \rp\ colour, the apparent magnitude, and the absolute magnitude.
For example, the mean radial velocity jitter of single stars with $\teff \approx 6000\,\textrm{K}$
will be higher than the jitter of single stars with $\teff \approx 5000\,\textrm{K}$, 
with all else being (nearly) equal. We similarly assume that the mean 
jitter of stellar \emph{multiples} will depend on the source colour and its position
in the Hertzsprung-Russell diagram. For these reasons, in order to evaluate whether a
source is more likely to be a single star or a stellar multiple, we must consider the 
jitter in context with other sources that have similar: \bp\ - \rp\
colour, apparent \rp\ magnitude, and absolute \rp\ magnitude. In principle if we knew
the distribution stellar multiple properties (or assumed one) then we could model the
observability of binary (and higher order) systems across the Hertszprung-Russell
diagram, but this is beyond the scope of this work.

% ADDITIONAL ASSUMPTIONS.
\todo{additional assumptions}

Given these assumptions, we chose to adopt a non-parametric model for stellar multiplicity,
primarily because we do not know the distribution of properties of stellar multiples.
Our approach is illustrated in Figure \ref{fig:npm_schematic} and can be qualitatively 
described by the following steps.\\

\noindent{}For each \Gaia\ source:
\begin{enumerate}

	\item Select $N$ sources that have a similar \bp\ - \rp\ colour, apparent \rp\
		  magnitude, and absolute \rp\ magnitude.

	\item Fit the distribution of radial velocity jitter for sources selected in
		  Step 1 using a mixture of a normal and a log-normal distribution.

	\item Evaluate whether the source of interest is more consistent with being
		  a single star component (drawn from the normal distribution) or a
		  stellar multiple (drawn from the log-normal distribution).

	\item If the source is more likely to be a stellar multiple, use the estimated
		  properties of single stars with similar properties to estimate the
		  excess radial velocity jitter due to the stellar companion(s).

@@

@@

% todo: joining sentence(s) here?

\subsection*{Step 1: Selecting similar sources}

We construct a $k$-dimensional tree \citep[k-d tree;][]{someone} for all sources in the
dimensions of \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude.
When doing so we scale the dimensions such that \todo{0.1}\,mags in \bp\ - \rp\ colour is
approximately equal to \todo{1} magnitude in apparent \rp\ magnitude and \todo{1} magnitude in 
absolute \rp\ magnitude. We use this $k$-d tree for selecting a `ball' of similar
stars for a given source.

We enforce a number of constraints when selecting similar stars. For each source
we require the multi-dimensional `ball' to include at least \todo{128} sources.
We further require that the radius of the ball extends to at least \todo{0.05}\,mags in 
\bp\ - \rp\ colour, \todo{0.5}\,mags in apparent \rp\ magnitude, and \todo{0.5}\,mags in absolute
\rp\ magnitude. If the initial query of \todo{128} points does not extend wide enough to
meet our constraints on minimum radius, we extend the radius (linearly in each scaled)
dimension until the minimum constraints on radii are met. Our minimum radii constraints
imply that there will be many (e.g., $>>\todo{128}$) points returned in dense regions of the
Hertszprung-Russell diagram. In these situations we then subsample the selected points
to return a random \todo{1024} sources in any ball, while enforcing our constraints on
minimum radii. As a result, the number of `similar' sources selected will range from
a minimum of \todo{128} to a maximum of \todo{1024}, and the minimum radii constraints on various
dimensions are always maintained.

% maximum ball radius

\subsection*{Step 2: Fitting the two-component mixture}

We fit a two-component mixture model to the distribution of radial velocity jitter for
all similar sources selected through Step 1. This mixture model includes a normal
distribution to represent the intrinsic jitter for single stars, and a log-normal
distribution to represent stellar multiples. Specifically the model parameters include
a mixing weight parameter $\theta$ that is constrained between $(0, 1)$, the mean 
$\mu_{s}$ and standard deviation $\sigma_{s}$ of the normal distribution,
and the mean $\mu_{m}$ and standard deviation $\sigma_{m}$ of the 
log-normal distribution.

We assume a beta prior on $\theta$  such that $\theta \sim \mathcal{B}\left(5, 5\right)$
and require that $\sigma_{s}$ and $\sigma_{m}$ are larger than zero.
Depending on the distribution of the radial velocity jitter, we often found this mixture
challenging to optimize because the log-normal distribution could easily account for 
most of the observed distribution, which tended the mixing weight to zero and implied
that all selected sources were stellar multiples. To avoid this situation we placed a
constraint on the mean of the log-normal distribution to ensure that the \emph{mode}
of the log-normal distribution is larger than $\mu_{s} + 1\sigma_{s}$ such that
\begin{eqnarray}
%	\mu_{s} + 5\sigma_{s} & > & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_lower_bound} \\
	\mu_{s} + 1\sigma_{s} & < & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_upper_bound} \quad .
\end{eqnarray}

\begin{comment}
However, we found that this constraint alone was insufficient. Often the mode would
be very close to the $\mu_{s} + 1\sigma_{s}$ lower bound, which implies that
the log-normal distribution is providing a considerable fraction of probability support
to sources with radial velocity jitter near the mean of the \emph{normal} distribution.
For this reason we chose to alter our lower bound constraint in that we also required
that at the mean of the normal distribution, the cumulative probability distribution
of the log-normal must be less than 10\% (let $q = 0.10$). The cumulative distribution 
function for the log-normal distribution is
\begin{eqnarray}
\textrm{CDF}(x|\mu_{m},\sigma_{m}) = \frac{1}{2} + \frac{1}{2}\textrm{erf}\left(\frac{\log{x} - \mu_{m}}{\sqrt{2}\sigma_{m}}\right) 
\end{eqnarray}

@@

We found these priors were sufficient to ensure numerical stability and efficiency.
\end{comment}

Wherever possible we initialised the model parameters from nearby points (within the
ball selected by the k-d tree) where an optimised result already exists. This was primarily to minimise
computational cost; the initialisation did not have an impact on the result. 
We required the log probability and the gradient of the log
probability to reach machine precision before considering the optimisation
converged. After convergence, we queried whether any points within the ball of
similar stars had not been optimized. If no result existed, we would next move to that
point. We refer to this process as \emph{swarm optimisation}, which we conducted in 
parallel, and illustrate in Figure \ref{fig:swarm_optimisation}.

@@

\begin{figure*}
%	\includegraphics[width=1.0\textwidth]{../figures/sb2_rvs_completeness.pdf}
	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
    \caption{Fraction of \Gaia\ sources with reported radial velocities
		     as a function of source properties. Within the adopted source
		     parameter ranges (gray; see Section \ref{sec:sb2_methods})
		     the radial velocity completeness is approximately flat.
		     We flag sources within this range that do not have reported 
		     radial velocities to be candidate SB2 systems.}
    \label{fig:sb2_rvs_completeness}
\end{figure*}

\begin{figure}
%	\includegraphics[width=0.4\textwidth]{../figures/sb2_histograms.pdf}
	\includegraphics[width=0.4\textwidth]{../figures/todo.png}
    \caption{Distribution of 
    			(a) astrometric unit weight error (see Eq. \ref{eq:astrometric_unit_weight_error}),
			%$u = \left(\chi_{al}^2/(N_{al} - 5)\right)^{1/2}$
			%		where $\chi_{al}^2$ and $N_{al}$ is the astrometric goodness-of-fit and number 
			%		of observations in the along-scan direction, respectively (columns 
			%			\texttt{astrometric\_chi2\_al} and \texttt{astrometric\_n\_obs\_al}), 
				(b) photometric \bp\ - \rp\ excess factor, 				
	 				and 
				(c) photometric \rp\ variability (see Eq. \ref{eq:photometric_variability})
			of candidate SB2 type systems relative to a control sample of the same size.}
    \label{fig:sb2_histograms}
\end{figure}

\begin{figure*}
%	\includegraphics[width=1.0\textwidth]{../figures/sb2_sky_structure.pdf}
	\includegraphics[width=1.0\textwidth]{../figures/todo.png}

    \caption{Fraction of sources (per sky bin) without reported radial velocities
    		 for all sources with $G \lesssim 13$ (top) and sources in the ranges
		 	 specified by Eqs \ref{eq:sb2_ranges} (bottom), where we assert that a missing
			 radial velocity measurement signifies a likely SB2 candidate . The color scale is 
			 arbitrarily set to highlight structure, where black indicates a higher
			 fraction of sources do not have reported radial velocities. Crowding
			 in the galactic plane likely results in some sources not having radial
			 velocities reported. The large scale structure visible in both axes is
			 a combined effect of the initial \Gaia\ source list, the scanning law,
			 and star forming regions (i.e., where emission in the \ion{Ca}{2} 
			 triplet likely causes issues for radial velocity determination).}
    \label{fig:sb2_sky_structure}
\end{figure*}

%\subsection{Photometric binaries} \label{sec:method_photometric_binaries}

%\begin{eqnarray}
%	\textrm{Amp} = \log_{10}\left(\sqrt{N_{obs,al}}\frac{\sigma_{\langle{}f_G\rangle}}{\langle{}f_G\rangle}\right) \label{eq:photometric_variability}
%\end{eqnarray}

\section{Results} \label{sec:results}
